Prospects of detecting gamma-ray emission from galaxy clusters: cosmic rays and dark matter annihilations

### 2011-06-13 paper.tex
update
updated fluxes, and XCR

We continue by comparing the estimated differential flux from Fornax
to three other clusters in Fig.~\ref{fig:clu_comp}; the close-by and
well studied Virgo cluster, the X-ray bright Perseus clusters, and the
massive merging Coma cluster. We find high DM induced gamma-ray fluxes
from the Fornax and Virgo clusters, which highlight them as promising
targets for indirect DM searches. \C{Note, however, that the angular
  extended of Virgo is 7 degrees, while the extended upper limit from
  Fermi-LAT is calculated assuming a $\sigma=1.2$ degree extension. In
  fact, the ratio of the virial radius to the assumed extension for
  the upper limits $\rvir/\sigma$ for $($Fornax, Coma, Virgo$)$ are
  $(2.79,\,1.3,\,5.8)$, respectively. Similarly, it is interesting to
  compare the radius that contain 68\% of the flux to the assumed
  extension, $r_\sigma/\sigma$, which is given by $(,\,,\,)$ for
  $($Fornax,Coma,Virgo$)$, respectively. It is quite striking how
  constraining the upper limits are for some of the clusters.}
\del{Note, however, that we have used spatially unresolved upper
  limits, which is quite optimistic for the Virgo cluster which has an
  angular extended of 7 degrees.  It is quite striking how
  constraining the upper limits are for the Fornax cluster are
  compared to the other clusters, especially in the $1-10$~GeV energy
  regime.} At these energies, Fermi-LAT has its peak sensitivity due
to a combination of increasing effective area and decreasing source
spectra as a function of energy. The upper limits for Virgo and
Perseus are unfortunately background contaminated by AGN activity from
M87 and NGC 1267, respectively, and do not gain much from the
increased sensitivity. Note the high CR-induced flux in Coma and
Perseus \C{is making it difficult for indirect DM searches in
  those clusters if the boost factors are much lower than assumed in
  this work.} \del{is of the same magnitude as the DM flux from LP
  models, making it difficult for indirect DM searches in those
  clusters.} Conversely, the Fornax cluster is a great target for
indirect DM studies because the relative low gamma-ray flux from CRs,
absents of an active AGN, and high DM gamma-ray flux. \del{The
  Fermi-LAT three-year data will be able to constrain the smallest
  mass of substructures using the DM BM models.}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.99\columnwidth]{figures/XCR.Fermi.eps}
  \caption{\it Limits on the relative CR pressure, $\bra X_\rmn{CR}\ket = \bra
    P_\rmn{CR} \ket / \bra P_\rmn{th}\ket$ averaged across the cluster. Our
    constraints are obtained with the extended Fermi-LAT limits in the energy
    range $E_\gamma=1-10$~GeV that assume a King profile which matches the
    extension of the thermal X-ray emission (black) and for comparison with
    Fermi-LAT point-source limits (grey) \cite{2010ApJ...717L..71A}. In
    computing those limits, we adopted our (conservative) analytic model
    \cite{2010MNRAS.409..449P} and a averaged relative CR flux as obtained from
    our simulations.}
 \label{fig:XCR}
\end{minipage}
\end{figure*}

@@

>>>>

We continue by comparing the estimated differential flux from Fornax
to three other clusters in Fig.~\ref{fig:clu_comp}; the close-by and
well studied Virgo cluster, the X-ray bright Perseus clusters, and the
massive merging Coma cluster. We find high DM induced gamma-ray fluxes
from the Fornax and Virgo clusters, which highlight them as promising
targets for indirect DM searches. \C{Note, however, that the angular
  extended of Virgo is 7 degrees, while the extended upper limit from
  Fermi-LAT is calculated assuming a $\sigma=1.2$ degree extension. In
  fact, the ratio of the virial radius to the assumed extension for
  the upper limits $\rvir/\sigma$ for $($Fornax, Coma, Virgo$)$ are
  $(2.8,\,1.6,\,5.3)$, respectively. Similarly, it is interesting to
  compare the radius that contain 68\% of the flux to the assumed
  extension, $r_\sigma/\sigma$, which is given by $(1.7,\,1.0,\,3.3)$ for
  $($Fornax,Coma,Virgo$)$, respectively. It is quite striking how
  constraining the upper limits are for some of the clusters.}
\del{Note, however, that we have used spatially unresolved upper
  limits, which is quite optimistic for the Virgo cluster which has an
  angular extended of 7 degrees.  It is quite striking how
  constraining the upper limits are for the Fornax cluster are
  compared to the other clusters, especially in the $1-10$~GeV energy
  regime.} At these energies, Fermi-LAT has its peak sensitivity due
to a combination of increasing effective area and decreasing source
spectra as a function of energy. The upper limits for Virgo and
Perseus are unfortunately background contaminated by AGN activity from
M87 and NGC 1267, respectively, and do not gain much from the
increased sensitivity. Note the high CR-induced flux in Coma and
Perseus \C{is making it difficult for indirect DM searches in
  those clusters if the boost factors are much lower than assumed in
  this work.} \del{is of the same magnitude as the DM flux from LP
  models, making it difficult for indirect DM searches in those
  clusters.} Conversely, the Fornax cluster is a great target for
indirect DM studies because the relative low gamma-ray flux from CRs,
absents of an active AGN, and high DM gamma-ray flux. \del{The
  Fermi-LAT three-year data will be able to constrain the smallest
  mass of substructures using the DM BM models.}

@@

@@

\C{A quantity that is of great scientific interest is the CR pressure
  ($P_\CR$) relative to the thermal pressure ($P_\rmn{th}$), $X_\CR =
  P_\CR/P_\rmn{th}$, as it directly bias galaxy cluster observables
  and the resulting cosmological constraints, such as the hydrostatic
  mass bias or the Sunyaev-Zel’dovich signal measured by Planck, South
  Pole Telescope, or Atacama Cosmology Telescope. In the following we
  estimate the mass scaling of the relative CR pressure using the CR
  formalism outlined in
  \cite{2007A&A...473...41E,2010MNRAS.409..449P}.  First we calculate
  the CR distribution function $f = C\,p^{-\alpha}$, where $C$ is the
  normalization of each power-law, $p=P/m_\rmn{p} c$ is the normalized
  CR proton momentum, and $\alpha$ is the spectral index. The CR
  pressure $P_\CR \propto C=\tilde{C}/\rho_\rmn{gas}$, where
  $\tilde{C}\propto M^{0.44}$ for clusters with a mass $\mvir\gtrsim
  10^{14}\,\msun$. We determine the mass scaling of the gas density
  through its relation to the enclosed gas fraction $f_\rmn{gas}$, and
  find that $\rho_\rmn{gas}\propto f_\rmn{gas} \propto M^{0.135}$
  \cite{2009ApJ...693.1142S}. The thermal pressure is derived from the
  temperature in virial equilibrium, and scales as $P_\rmn{th}\propto
  k_\B T \propto G M/R \propto M^{0.67}$. Finally we can derive the
  mass scaling of the relative CR pressure from,
\begin{equation}
 X_\CR = \frac{P_\CR}{P_\rmn{th}} \propto \frac{\rho_\rmn{gas}\,
   \tilde{C}}{k_\B T} \propto M^{-0.06}\,,
\end{equation}
which is roughly a constant, i.e. almost independent on mass. Since
the CRs that build up the bulk of the CR pressure also give rise to
gamma-ray emission through the production of decaying pions, we can
use gamma-ray upper limits from Fermi-LAT to constrain $X_\CR$. These
upper limits are 95 percent confidence limits and assume a spectral
index of 2. EXPAND DISCUSSION, ADD VOLUME WEIGHT, HOW TO GO FROM
GAMMA->XCR.}

\C{In Fig.~\ref{fig:XCR} we show the volume weighted relative CR
  pressure for 32 clusters, for which Fermi-LAT gamma-ray upper limits
  are derived. We find that the most constraining clusters are the
  Norma and Coma clusters of $X_\CR < 0.07$.}

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.99\columnwidth]{figures/XCR.Fermi.eps}
  \caption{\it Limits on the relative CR pressure, $< X_\rmn{CR}> = <
    P_\rmn{CR} > / < P_\rmn{th}>$ averaged across the cluster. Our
    constraints are obtained with the extended Fermi-LAT limits in the energy
    range $E_\gamma=1-10$~GeV that assume a King profile which matches the
    extension of the thermal X-ray emission (black) and for comparison with
    Fermi-LAT point-source limits (grey) \cite{2010ApJ...717L..71A}. In
    computing those limits, we adopted our (conservative) analytic model
    \cite{2010MNRAS.409..449P} and an averaged relative CR flux as obtained from
    our simulations.}
 \label{fig:XCR}
\end{minipage}
\end{figure*}
